---
title: "Reproducible Research Assignment"
author: "David Contreras Carretero"
date: "5/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data

Data for this assignment was included in the initial repository in a .zip file named activity.zip containing a .csv file named activity.csv.

```{r echo=TRUE}
data <- read.table(unz("activity.zip", "activity.csv"),header =T , quote="\"", sep=",")
```

## What is mean total number of steps taken per day?


Frist I will summarise the data and obtain a data set with one entry for each day with total steps for each day.

```{r }
library(dplyr)
Total_Steps <- data %>% group_by(date) %>% summarise(steps=sum(steps) )
```

### Histogram of total number of steps taken each day

```{r}
hist(Total_Steps$steps, xlab="Steps",main="Total steps in one day")
```


### Median and Mean of total steps per day

```{r }
steps_mean <- mean(Total_Steps[!is.na(Total_Steps$steps),]$steps)
steps_median <- median(Total_Steps[!is.na(Total_Steps$steps),]$steps)
```

The mean of steps per day is `r steps_mean`, and the median `r steps_median`.


## What is the average daily activity pattern?

Frist I calculate the mean of steps for each interval across all dates. So I group by interval and summarise using mean function, but using only the complete cases, that is ignoring missing values.

```{r}
DailyActivity <- data[complete.cases(data),] %>% group_by(interval) %>% summarise(StepsMean=mean(steps) )
```

And then plot it.

```{r}
plot(DailyActivity$interval, DailyActivity$StepsMean,type="l",xlab="Interval", ylab="Steps")
```





