---
title: "Reproducible Research: Peer Assessment 1"
author: "David Contreras Carretero"
date: "5/9/2020"
output: html_document
html_document:
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options("scipen"=100, "digits"=2)
Sys.setlocale("LC_ALL","English")
```

## Loading and preprocessing the data
Data for this assignment was included in the initial repository in a .zip file named activity.zip containing a .csv file named activity.csv.

Then I convert variable "date", currently a character string to a date value.  

```{r echo=TRUE}
data <- read.table(unz("activity.zip", "activity.csv"),header =T , quote="\"", sep=",")
library(dplyr)
data <- data %>% mutate(date = as.Date(date))
```



## What is mean total number of steps taken per day?

Frist I will summarise the data and obtain a data set with one entry for each day with total steps for each day.

```{r }
Total_Steps <- data %>% group_by(date) %>% summarise(steps=sum(steps) )
```


```{r}
hist(Total_Steps$steps, xlab="Steps",main="Total steps in one day")
```

```{r }
steps_mean <- mean(Total_Steps[!is.na(Total_Steps$steps),]$steps)
steps_median <- median(Total_Steps[!is.na(Total_Steps$steps),]$steps)

```

The mean of steps per day is `r steps_mean`, and the median `r steps_median`.


## What is the average daily activity pattern?

Frist I calculate the mean of steps for each interval across all dates. So I group by interval and summarise using mean function, but using only the complete cases, that is ignoring missing values.

```{r}
DailyActivity <- data[complete.cases(data),] %>% group_by(interval) %>% summarise(StepsMean=mean(steps) )
```

And then plot it.

```{r}
plot(DailyActivity$interval, DailyActivity$StepsMean,type="l",xlab="Interval", ylab="Steps",main="Average Daily Activity Pattern")
```


## Imputing missing values

Frist I check how many missing values are there and in which column.

``` {r}
colSums(is.na(data))
```

There are 2304 missing values, only in "steps" column.

I'll fill those missing values with the mean of the interval for the rest of the days. I already have calculated the mean of each interval for the previous plot, I just need to update the missing values with their corresponding interval mean value.

```{r}
data <- data %>% left_join(DailyActivity, by="interval") %>%
mutate(steps= coalesce(steps, StepsMean ) ) %>%
select(-StepsMean)
```

Now i'll repeat the firs histogram, with  total number of steps taken each day.


```{r}
Total_Steps <- data %>% group_by(date) %>% summarise(steps=sum(steps) )
hist(Total_Steps$steps, xlab="Steps",main="Total steps in one day")
```

```{r }
steps_mean <- mean(Total_Steps[!is.na(Total_Steps$steps),]$steps)
steps_median <- median(Total_Steps[!is.na(Total_Steps$steps),]$steps)
```

The new of mean of steps per day is `r steps_mean`, and the median `r steps_median`.

Mean is exactly the same as previous calculation without estimated values, which make sense because be used the mean of the interval for filling missing values. The median increases a little bit and matches exactly the mean. 

In the histogram we see an increase of the middle bar area (10000 - 15000), previously had a frecuency of around 30 and now has raised until 35.

## Are there differences in activity patterns between weekdays and weekends?

First I create the factor weekend indicating wherter the date is a "weekday" or a "weekend" day. 

```{r}
data <- data %>% mutate(weekday = weekdays(date),weekend =ifelse(weekday %in% c("Saturday","Sunday"),"weekend","weekday"))
data$weekend <- factor(data$weekend)
```
Then  I summarise the mean of steps for weekdays and weekend in two diffent data frames.
```{r}
Activity <- data %>% group_by(weekend,interval) %>% summarise(StepsMean=mean(steps) )
ActivityWeekday <- Activity[Activity$weekend =="weekday",] 
ActivityWeekend <- Activity[Activity$weekend =="weekend",] 

```

```{r}
par(mfrow=c(2,1),pty="m",mar=c(1,1,1,0), oma=c(1,3,2,2))

plot(ActivityWeekday$interval,ActivityWeekday$StepsMean,type="l",main="Activity pattern weekdays vs. weekends",bty="n", xlim=c(0,2355), xaxt='n')
mtext("Weekdays",  side=4)
mtext("Steps",  side=2)
plot(ActivityWeekend$interval,ActivityWeekend$StepsMean,type="l",bty="n",xlim=c(0,2355),xaxt='n')
axis(side =1, at=seq(0, 2400, by = 100))
mtext("Weekdends",  side=4)
mtext("Interval",  side=1)
mtext("Steps",  side=2)

```

It seems that test subject had more activity along the day during weekends in comparison with weekdays where the activity is more concentrated during the first intervals of the day.




